---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# OMOPGenerics

[![Lifecycle:Experimental](https://img.shields.io/badge/Lifecycle-Experimental-339999)](https://lifecycle.r-lib.org/articles/stages.html)

The OMOPGenerics package provides definitions of core classes and methods used by analytic pipelines that query the OMOP common data model.

## Installation

You can install the development version of OMOPGenerics from [GitHub](https://github.com/) with:

```{r, eval=FALSE}
install.packages("remotes")
devtools::install_github("darwin-eu-dev/OMOPGenerics")
```

And load it using the library command:
```{r}
library(OMOPGenerics)
```


## Core classes and methods

### CDM Reference

A cdm reference is a single R object that represents OMOP CDM data. The tables in the cdm reference may be in a database, but a cdm reference may also contain OMOP CDM tables that are in dataframes/tibbles or in arrow. In the latter case the cdm reference would typically be a subset of an original cdm reference that has been derived as part of a particular analysis.

OMOPGenerics contains the class definition of a cdm reference and a dataframe implementation. For creating a cdm reference using a database, see the CDMConnector package (<https://darwin-eu.github.io/CDMConnector/>).

A cdm object can contain three type of tables:

* Standard tables:
```{r}
omopTables()
```
Each one of the tables has a required columns. For example, for the `person` table this are the required columns:
```{r}
omopColumns(table = "person")
```

* Cohort tables (see `generatedCohortSet`).

* Other tables, these other tables can have any format. 

Any table to be part of a cdm object has to fulfill 4 conditions:

- All must share a common source.

- The name of the tables must be lowercase.

- The name of the column names of each table must be lowercase.

- `person` and `observation_period` must be present.


### Concept set

A concept set can be represented as either a codelist or a concept set expression. A codelist is a named list, with each item of the list containing specific concept IDs. 
```{r}
condition_codes <- list("diabetes" = c(201820, 4087682, 3655269),
                        "asthma" = 317009)
condition_codes <- codelist(condition_codes)

condition_codes
```

Meanwhile, a concept set expression provides a high-level definition of concepts that, when applied to a specific OMOP CDM vocabulary version (by making use of the concept hierarchies and relationships), will result in a codelist. 

```{r}
condition_cs <- list(
  "diabetes" = dplyr::tibble(
    "concept_id" = c(201820, 4087682),
    "excluded" = c(FALSE, FALSE),
    "descendants" = c(TRUE, FALSE),
    "mapped" = c(FALSE, FALSE)
  ),
  "asthma" = dplyr::tibble(
    "concept_id" = 317009,
    "excluded" = FALSE,
    "descendants" = FALSE,
    "mapped" = FALSE
  )
)
condition_cs <- conceptSetExpression(condition_cs)

condition_cs
```

### Generatred cohort set


### Summarised result


### Compared result


### Participants
