---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# omopgenerics

<!-- badges: start -->
[![Lifecycle:Experimental](https://img.shields.io/badge/Lifecycle-Experimental-339999)](https://lifecycle.r-lib.org/articles/stages.html)
[![R-CMD-check](https://github.com/darwin-eu-dev/omopgenerics/workflows/R-CMD-check/badge.svg)](https://github.com/darwin-eu-dev/omopgenerics/actions)
[![CRANstatus](https://www.r-pkg.org/badges/version/omopgenerics)](https://CRAN.R-project.org/package=omopgenerics)
[![Codecov test coverage](https://codecov.io/gh/darwin-eu-dev/omopgenerics/branch/main/graph/badge.svg)](https://app.codecov.io/gh/darwin-eu-dev/omopgenerics?branch=main)
<!-- badges: end -->

## Package overview

The omopgenerics package provides definitions of core classes and
methods used by analytic pipelines that query the OMOP common data
model.

``` {r, echo =FALSE}
citation("omopgenerics")
```

If you find the package useful in supporting your research study, please
consider citing this package.

## Installation

You can install the development version of OMOPGenerics from [GitHub](https://github.com/) with:

```{r, eval=FALSE}
install.packages("remotes")
devtools::install_github("darwin-eu-dev/omopgenerics")
```

And load it using the library command:
```{r, message=FALSE, warning=FALSE}
library(omopgenerics)
library(dplyr)
```


## Core classes and methods

### CDM Reference

A cdm reference is a single R object that represents OMOP CDM data. The tables in the cdm reference may be in a database, but a cdm reference may also contain OMOP CDM tables that are in dataframes/tibbles or in arrow. In the latter case the cdm reference would typically be a subset of an original cdm reference that has been derived as part of a particular analysis.

omopgenerics contains the class definition of a cdm reference and a dataframe implementation. For creating a cdm reference using a database, see the CDMConnector package (<https://darwin-eu.github.io/CDMConnector/>).

A cdm object can contain four type of tables:

* Standard tables:
```{r}
omopTables()
```
Each one of the tables has a required columns. For example, for the `person` table this are the required columns:
```{r}
omopColumns(table = "person")
```

* Cohort tables
We can see the cohort-related tables and their required columns.
```{r}
cohortTables()
cohortColumns(table = "cohort")
```

In addition, cohorts are defined in terms of a `generatedCohortSet` class. For more details on this class definition see the corresponding vignette.

* Achilles tables 
The Achilles R package generates descriptive statistics about the data contained in the OMOP CDM. Again, we can see the tables created and their required columns. 
```{r}
achillesTables()
achillesColumns(table = "achilles_results")
```

* Other tables, these other tables can have any format. 

Any table to be part of a cdm object has to fulfill 4 conditions:

- All must share a common source.

- The name of the tables must be lowercase.

- The name of the column names of each table must be lowercase.

- `person` and `observation_period` must be present.


### Concept set

A concept set can be represented as either a codelist or a concept set expression. A codelist is a named list, with each item of the list containing specific concept IDs. 
```{r}
condition_codes <- list("diabetes" = c(201820, 4087682, 3655269),
                        "asthma" = 317009)
condition_codes <- newCodelist(condition_codes)

condition_codes
```

Meanwhile, a concept set expression provides a high-level definition of concepts that, when applied to a specific OMOP CDM vocabulary version (by making use of the concept hierarchies and relationships), will result in a codelist. 

```{r}
condition_cs <- list(
  "diabetes" = dplyr::tibble(
    "concept_id" = c(201820, 4087682),
    "excluded" = c(FALSE, FALSE),
    "descendants" = c(TRUE, FALSE),
    "mapped" = c(FALSE, FALSE)
  ),
  "asthma" = dplyr::tibble(
    "concept_id" = 317009,
    "excluded" = FALSE,
    "descendants" = FALSE,
    "mapped" = FALSE
  )
)
condition_cs <- newConceptSetExpression(condition_cs)

condition_cs
```

### A cohort table

A cohort is a set of persons who satisfy one or more inclusion criteria for a duration of time and, when defined, this table in a cdm reference has a cohort table class. Cohort tables are then associated with attributes such as settings and attrition. 

```{r}
person <- tibble(
  person_id = 1, gender_concept_id = 0, year_of_birth = 1990,
  race_concept_id = 0, ethnicity_concept_id = 0
)
observation_period <- dplyr::tibble(
  observation_period_id = 1, person_id = 1,
  observation_period_start_date = as.Date("2000-01-01"),
  observation_period_end_date = as.Date("2023-12-31"),
  period_type_concept_id = 0
)
diabetes <- tibble(
  cohort_definition_id = 1, subject_id = 1,
  cohort_start_date = as.Date("2020-01-01"),
  cohort_end_date = as.Date("2020-01-10")
)

cdm <- cdmFromTables(
  tables = list(
    "person" = person,
    "observation_period" = observation_period,
    "diabetes" = diabetes
  ),
  cdmName = "example_cdm"
)
cdm$diabetes <- newCohortTable(cdm$diabetes)

cdm$diabetes
settings(cdm$diabetes)
attrition(cdm$diabetes)
cohortCount(cdm$diabetes)
```

### Summarised result

A summarised result provides a standard format for the results of an analysis performed against data mapped to the OMOP CDM.

For example this format is used when we get a summary of the cdm as a whole
```{r}
summary(cdm) |> 
  dplyr::glimpse()
```

and also when we summarise a cohort
```{r}
summary(cdm$diabetes) |> 
  dplyr::glimpse()
```


