---
title: "a01_cdm_reference"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{a01_cdm_reference}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(OMOPGenerics)
```

## The cdm reference

A cdm reference is a single R object that represents OMOP CDM data. The tables in the cdm reference may be in a database, but a cdm reference may also contain OMOP CDM tables that are in dataframes/tibbles or in arrow. In the latter case the cdm reference would typically be a subset of an original cdm reference that has been derived as part of a particular analysis.

OMOPGenerics contains the class definition of a cdm reference and a dataframe implementation. For creating a cdm reference using a database, see the CDMConnector package (<https://darwin-eu.github.io/CDMConnector/>).

A cdm object can contain three type of tables:

* Standard tables:
```{r}
standardOmopCdmTables()
```
Each one of the tables has a required columns. For example, for the `person` table this are the required columns:
```{r}
requiredTableColumns(table = "person")
```

* Cohort tables (see `generatedCohortSet`).

* Other tables, these other tables can have any format. 

Any table to be part of a cdm object has to fulfill 4 conditions:

- All must share a common source.

- The name of the tables must be lowercase.

- The name of the column names of each table must be lowercase.

- `person` and `observation_period` must be present.

### Required columns

```{r, echo = FALSE}
version = "5.3"
mes <- cli::col_green(cli::style_bold("OMOP CDM version: \"{version}\""))
for (tab in c("person", "observation_period")) {
  mes <- c(
    mes, 
    "*" = paste0(
      cli::col_blue(cli::style_bold(tab)), ": ", paste0(
        requiredTableColumns(table = tab, version = version), collapse = ", "
      )
    )
  )
}
tables <- standardOmopCdmTables(version = version)
tables <- sort(tables[!tables %in% c("person", "observation_period")])
for (tab in tables) {
  mes <- c(
    mes, 
    "*" = paste0(
      cli::col_blue(tab), ": ", paste0(
        requiredTableColumns(table = tab, version = version), collapse = ", "
      )
    )
  )
}
cli::cli_inform(mes)
```

```{r, echo = FALSE}
version = "5.4"
mes <- cli::col_green(cli::style_bold("OMOP CDM version: \"{version}\""))
for (tab in c("person", "observation_period")) {
  mes <- c(
    mes, 
    "*" = paste0(
      cli::col_blue(cli::style_bold(tab)), ": ", paste0(
        requiredTableColumns(table = tab, version = version), collapse = ", "
      )
    )
  )
}
tables <- standardOmopCdmTables(version = version)
tables <- sort(tables[!tables %in% c("person", "observation_period")])
for (tab in tables) {
  mes <- c(
    mes, 
    "*" = paste0(
      cli::col_blue(tab), ": ", paste0(
        requiredTableColumns(table = tab, version = version), collapse = ", "
      )
    )
  )
}
cli::cli_inform(mes)
```

## Export the cdm reference

When the export method is applied to a cdm reference, metadata about that cdm will be written to a csv. The csv contains the following columns

| Variable | Description | 
|------|------|
|result_type|      |
|cdm_name|      |
|cdm_source_name|      |
|cdm_description|      |
|cdm_documentation_reference|      |
|cdm_version|      |
|cdm_holder|      |
|cdm_release_date|      |
|vocabulary_version|      |
|person_count|      |
|observation_period_count|      |
|earliest_observation_period_start_date|      |
|latest_observation_period_end_date|      |
|snapshot_date|      |

export method ....
