---
title: "a01_cdm_reference"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{a01_cdm_reference}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(OMOPGenerics)
```

A cdm reference is a single R object that represents OMOP CDM data. The tables in the cdm reference may be in a database, but a cdm reference may also contain OMOP CDM tables that are in dataframes/ tibbles or in arrow. In the latter case the cdm reference would typically be a subset of an original cdm reference that has been derived as part of a particular analysis.

OMOPGenerics contains the class definition of a cdm reference and a dataframe implementation. For creating a cdm reference using a database, see the CDMConnector package (<https://darwin-eu.github.io/CDMConnector/>).

```{r}
OMOPGenerics::cdmReference
```

When the export method is applied to a cdm reference, metadata about that cdm will be written to a csv. The csv contains the following columns

| Variable | Description | 
|------|------|
|result_type|      |
|cdm_name|      |
|cdm_source_name|      |
|cdm_description|      |
|cdm_documentation_reference|      |
|cdm_version|      |
|cdm_holder|      |
|cdm_release_date|      |
|vocabulary_version|      |
|person_count|      |
|observation_period_count|      |
|earliest_observation_period_start_date|      |
|latest_observation_period_end_date|      |
|snapshot_date|      |

export method ....
