---
title: "Validate a cohort_table"
author: "Martí Català"
date: "`r Sys.Date()`"

---

## Cohort table

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(omopgenerics)
```


A cohort is a **set of people that fulfill a certain set of criteria for a period of time**.

In omopgenerics we defined the `cohort_table` class that allows us to represent individuals in a cohort.

A `cohort_table` is created using the `newCohortTable()` function that is defined by:

- A cohort table.

- A cohort set.

- A cohort attrition.

The cohort set and the cohort attrition will be instantiated in the database every time that we create a new cohort table. You can access the cohort set of a cohort_table using the function `settings`:
```{r, eval = FALSE}
settings(cdm$cohort)
```

You can access the cohort attrition of a cohort_table using the function `attrition`:
```{r, eval = FALSE}
attrition(cdm$cohort)
```

Each one of the elements have to fulfill certain criteria.

### Cohort Set

A cohort set must be a table with:

- Lower case column names.

- At least `r cohortColumns("cohort_set")` columns (`cohortColumns("cohort_set")`).

- `cohort_name` it must contain unique cohort names (currently they are cased to snake case).

- `cohort_definition_id` it must contain unique cohort ids, all the ids present in table must be present in the cohort set and the same ids must be present in cohort attrition.

### Cohort Attrition

A cohort attrition must be a table with:

- Lower case column names.

- At least `r cohortColumns("cohort_attrition")` columns (`cohortColumns("cohort_attrition")`).

- `cohort_definition_id` it must contain cohort ids, all the ids present in table must be present in the cohort attrition and the same ids must be present in cohort set.

- There must exist unique pairs of `cohort_definition_id` and `reason_id`.

### Cohort Table

A cohort table must be a table with:

- It comes from a cdm_reference (extracted via `cdm$cohort`).

- It has the same source than this cdm_reference.

- Lower case column names.

- At least `r cohortColumns("cohort")` columns (`cohortColumns("cohort")`).

- There is no record with `NA` value in the required columns.

- There is no record with `cohort_start_date` > `cohort_end_date`.

- There is no overlap between observations. A person can be in a cohort several times (several records with the same subject_id). But it can't enter the cohort again before leaving it. So an individual can't be simultaneously more than once in the same cohort. This rule is applied at the cohort_definition_id level, so records with different cohort_definition_id can overlap.

- All the time between cohort_start_date and cohort_end_date (both included) the individual must be in observation.
